---
title: "Radar Chart example"
author: "Min Hwangbo"
date: "6/14/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Acknowledegements

* Inpisration and steps for the current project comes from [Zach, 2020](https://www.statology.org/radar-chart-in-r/#:~:text=%20In%20order%20to%20create%20a%20radar%20chart,to%20display%20on%20the%20radar%20chart.%20More%20) @ [Statology](https://www.statology.org/).

* The data used in this practice are collceted via Leadership Practices Inventory (LPI) 360 created by Kouzes, J. & Posner, B. (1987) @ [The Leadership Challenge](https://www.leadershipchallenge.com/) - A Wiley Brand. 

* Five practices of exemplary leadership summary data are represented including 1) Model the way; 2) Inspire a shared vision; 3) Challenge the process; 4) Enable others to act; and, 5) Encourage the heart.

* The data used in this practice are my personal performance metrics reported by myself, two direct reports, five colleagues, and two supervisors (n=10)

# Load a package
```{r, message = F, warning = F}
library(fmsb)  # Medical metrics / demographic data package
library(readr) # CSV loading package
```

# Load a data set 
```{r}
LPIdata <- read_csv("~/GitHub/InnovationLab/PerformanceMetrics/LPI/Data/MH_LPI_2019.csv")
```

## Quality check
```{r}
head(LPIdata) # first five records
tail(LPIdata) # last five records
ls(LPIdata)   # vars in the data set
```